---
title: "Code Workshop 1"
author: "Yue Tian"
date: "4/3/2021"
output: html_document
---


## Introduction 

For the series of workshops we will cover data manipulation, data visualization and model implementation.
This workshop will help you have more confidence with processing data in R.

Today, we will start the first part, data manipulation. 
Data manipulation includes 

- Get the data you wanted
- Arrange the data
- Clean the data

We will use the following packages to help us. 

- dplyr
- tidyverse
- nycflights13(this is the data we will use)

## Installing packages

```{r}
#install.packages("dplyr")
#install.packages("tidyverse")
#install.packages("nycflights13")
```

## Call packages

```{r}
library("dplyr")
library("tidyverse")
library("nycflights13")
```

## Preview of data
Whenever we have the data, the first thing we need to do is to exam it. 
- This is the data of flights of new york airports in 2013
- The attributes includes the flight month, day and time, the destination, the airliner, the delay time and their flight hours.
- Using head() function only shows the first couple rows of the dataframe. This will help us save time when we are dealing with very large data

```{r}
dat = flights
head(dat)
```

## dplry basics 
1. Filter the rows with filter() function 

- filter() functions allows you to subset the data based on their values 
- for example, I want to find all the flights in January 1st
- For the filter() functions the first argument is the name of the data and other arguments are the conditions of the filter

```{r}
jan1 = filter(dat,month == 1,day ==1)
print(jan1)
```
## Logic operators 
At the Filter functions there is something like this "=="
- This is a logic operator meaning equal.
- In the filter function above, "month == 1" means that we want data that month is 1
- There are other logic operators
- ">" : larger than 
- "<" : smaller than
- "!=" : not equal
```{r}
2 == 1
```

```{r}
3!= 2
```


## Other logic operators

Other logic operator includes:
- "&": and (we don't need this in filter() as a comma in filter in filter() means and)
- "|": or
- For example we want to get the flight in January and February 
```{r}
filter(dat, month == 1 | month ==2)
```
## Exercise
========================================================
- Find the flights that has an arrival delay is more than 1 hour (the attribute name is: arr_delay)
```{r}
flight_delay = filter(dat,arr_delay >= 60)
```

- Find the all United flights flew to Houston (attribute name: dest, Houston airports : IAH or HOU)
```{r}
United_huston = filter(dat,carrier == "UA", dest == "IAH" | dest == "HOU")
```

## Arrange rows using arrange()

-arrange() works like filter() and except it changes the order of rows rather than selecting them

- if you want the data in descending order use desc()

- arrange the flights based on their delay (attribute arr_delay)
```{r}
arrange(dat,arr_delay)
```

- put them in a descending order

```{r}
arrange(dat,desc(arr_delay))
```

## Practice : find the flight with least time in air (attribute is air_time)
```{r}
arrange(dat,air_time)
```

## Select Columns using select()

- If you just want a few attributes from your dataframe simply use select() to get them
```{r}
select(dat,year: day)#(inclusive)
```

- you can also exclude the columns 
- Select all the columns except year day and month 
```{r}
select(dat,-(year:day))
```

- Other functions that helps select

- starts_with("abc")
- For example: we want attributes about arrival
```{r}
select(dat,starts_with("arr"))
```

- ends_with("abc") # this is similiar to starts_with()

-contains("abc")
-For example, we want to get information about delays
```{r}
select(dat,contains("delay"))
```

## Add new attributes with mutate()

- Sometimes we also want to add variables to our dataframe, for instance add an average. And we can use mutate() to add a new column 

- notice: mutate adds new columns at the end of your dataframe

- Before we start, let's make a narrow dataframe with less attributes
```{r}
delay_flights =  select(dat, year:day, ends_with("delay"), distance, air_time)

```


- if I want to calculate the speed of the delaied flights 
```{r}
delay_flights =  select(dat, year:day, ends_with("delay"), distance, air_time)
delay_flights_with_speed = mutate(delay_flights,speed = distance / air_time*60)
```

## Summaries using summarize()

- get the average delay for entire 2013 
- na.rm = TRUE : we remove all NA's in our dataframe 
```{r}
summarize(dat,delay = mean(dep_delay,na.rm = TRUE))
```


- Use group_by() with summarize
- Let's calculate average delay per day
```{r}
by_day = group_by(dat,year,month,day)
summarize(by_day,delay = mean(dep_delay,na.rm = TRUE))
```




